---
title: "STAT253 - Homework#1"
author: "Osman MANTICI - 150117505"
date: "29 Mart 2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r stat,echo=FALSE}

x<-read.table(file="/Users/osman/Documents/R/HW#1/HW1_Data_v1.csv", header=TRUE, sep=",")

attach(x)
```
1-)
  Systolic bp, and diastolic bp and age variables are multivariate, discrete and quantitative.
  The sex variable is qualitative, continuous, and bivariate.

2-)
  Dotplot, histogram, boxplot, stem-and-leaf plot, can be used when this data set read.
  
  When the questions are want to learn about the range, median and mean, maximum and minimum values, where is the collapses or outliers, answers are found easily. Furthermore, boxplot gives us about the five number summary rather than the others. Stem-and-leaf plot and dotplot also gives more  the exact information.  


3-)

a) Means of men and women

Results are in that order, 
men systolic blood pressure mean,
women systolic blood pressure mean, 
men diastolic blood pressure mean, 
women diastolic blood pressure mean. 
  
```{r a, echo=FALSE}
men_sys_mean <- mean(SYSBP[GENDER==0])
men_sys_mean

women_sys_mean <- mean(SYSBP[GENDER==1])
women_sys_mean


men_dias_mean <- mean(DIASBP[GENDER==0])
men_dias_mean

women_dias_mean <- mean(DIASBP[GENDER==1])
women_dias_mean
```


b) Sample variance of men and women

Results are in that order
  men systolic blood pressure sample variance
  women systolic blood pressure sample variance
  men diastolic blood pressure sample variance
  women diastolic blood pressure sample variance
  
```{r b,echo=FALSE}
men_sys_variance <- var(SYSBP[GENDER==0])
men_sys_variance

women_sys_variance <- var(SYSBP[GENDER==1])
women_sys_variance


men_dias_variance <- var(DIASBP[GENDER==0])
men_dias_variance

women_dias_variance <- var(DIASBP[GENDER==1])
women_dias_variance
```

c) Standard deviation of men and women

Results are in that order
  
  men systolic blood pressure standard deviation
  women systolic blood pressure standard deviation
  men diastolic blood pressure standard deviation
  women diastolic blood pressure standard deviation

```{r c,echo=FALSE}

men_sys_stddev <- sd(SYSBP[GENDER==0])
men_sys_stddev

women_sys_stddev <- sd(SYSBP[GENDER==1])
women_sys_stddev

men_dias_stddev <- sd(DIASBP[GENDER==0])
men_dias_stddev

women_dias_stddev <- sd(DIASBP[GENDER==1])
women_dias_stddev
```

d) 1. Lower quartiles of men and women

Results are in that order
  
  men systolic blood pressure lower quartiles
  women systolic blood pressure lower quartiles
  men diastolic blood pressure lower quartiles
  women diastolic blood pressure lower quartiles
  
```{r d1,echo=FALSE}
men_sys_lowerquartile <- quantile(SYSBP[GENDER==0], 0.25)
men_sys_lowerquartile

women_sys_lowerquartile <- quantile(SYSBP[GENDER==1], 0.25)
women_sys_lowerquartile


men_dias_lowerquartile <- quantile(DIASBP[GENDER==0], 0.25)
men_dias_lowerquartile

women_dias_lowerquartile <- quantile(DIASBP[GENDER==1], 0.25)
women_dias_lowerquartile
```


d) 2. Upper quartiles of men and women

Results are in that order
  men systolic blood pressure upper quartiles
  women systolic blood pressure upper quartiles
  men diastolic blood pressure upper quartiles
  women diastolic blood pressure upper quartiles
  
```{r d2,echo=FALSE}
men_sys_upperquartile <- quantile(SYSBP[GENDER==0], 0.75)
men_sys_upperquartile

women_sys_upperquartile <- quantile(SYSBP[GENDER==1], 0.75)
women_sys_upperquartile


men_dias_upperquartile <- quantile(DIASBP[GENDER==0], 0.75)
men_dias_upperquartile

women_dias_upperquartile <- quantile(DIASBP[GENDER==1], 0.75)
women_dias_upperquartile

```


e) 1. Maximum values of men and women

Results are in that order
  men systolic blood pressure maximum
  women systolic blood pressure maximum
  men diastolic blood pressure maximum
  women diastolic blood pressure maximum
  
  
```{r e1,echo=FALSE}
men_sys_max <- max(SYSBP[GENDER==0])
men_sys_max

women_sys_max <- max(SYSBP[GENDER==1])
women_sys_max


men_dias_max <- max(DIASBP[GENDER==0])
men_dias_max

women_dias_max <- max(DIASBP[GENDER==1])
women_dias_max
```

e) 2. Minimum values of men and women

Results are in that order
  men systolic blood pressure minimum
  women systolic blood pressure minimum
  men diastolic blood pressure minimum
  women diastolic blood pressure minimum
  
  
```{r e2,echo=FALSE}
men_sys_min <- min(SYSBP[GENDER==0])
men_sys_max

women_sys_min <- min(SYSBP[GENDER==1])
women_sys_min


men_dias_min <- min(DIASBP[GENDER==0])
men_dias_min

women_dias_min <- min(DIASBP[GENDER==1])
women_dias_min
```


f) Range values of men and women

Results are in that order
  men systolic blood pressure range
  women systolic blood pressure range
  men diastolic blood pressure range
  women diastolic blood pressure range
  

```{r f,echo=FALSE}
men_sys_range <- men_sys_max - men_sys_min
men_sys_range

women_sys_range <- women_sys_max - women_sys_min
women_sys_range


men_dias_range <- men_dias_max - men_dias_min
men_dias_range

women_dias_range <- women_dias_max - women_dias_min
women_dias_range
```


g) Comparing range and standart deviations, Range values how many standard deviation of men and women

Results are in that order
  men systolic blood pressure number of standart deviations 
  women systolic blood pressure number of standart deviations
  men diastolic blood pressure number of standart deviations
  women diastolic blood pressure number of standart deviations
  
  
```{r g,echo=FALSE}
men_sys_numberofstd <- men_sys_range/men_sys_stddev
men_sys_numberofstd

women_sys_numberofstd <- women_sys_range/women_sys_stddev
women_sys_numberofstd

men_dias_numberofstd <- men_dias_range/men_dias_stddev
men_dias_numberofstd

women_dias_numberofstd <- women_dias_range/women_dias_stddev
women_dias_numberofstd
```


h) Median values of men and women

Results are in that order
  men systolic blood pressure median
  women systolic blood pressure median
  men diastolic blood pressure median
  women diastolic pressure median
  
  
```{r h,echo=FALSE}
men_sys_median <- median(SYSBP[GENDER==0])
men_sys_median

women_sys_median <- median(SYSBP[GENDER==1])
women_sys_median


men_dias_median <- median(DIASBP[GENDER==0])
men_dias_median

women_dias_median <- median(DIASBP[GENDER==1])
women_dias_median
```

i) IQR's  of men and women

Results are in that order
  men systolic blood pressure IQR's
  women systolic blood pressure IQR's
  men diastolic blood pressure IQR's
  women diastolic blood pressure IQR's
  
  
```{r i,echo=FALSE}
men_sys_IQR <- IQR(SYSBP[GENDER==0])
men_sys_IQR

women_sys_IQR <- IQR(SYSBP[GENDER==1])
women_sys_IQR


men_dias_IQR <- IQR(DIASBP[GENDER==0])
men_dias_IQR

women_dias_IQR <- IQR(DIASBP[GENDER==1])
women_dias_IQR
```

j) 5-number-summary  of men and women

Results are in that order
  men systolic blood pressure 5-number-summary
  women systolic blood pressure 5-number-summary
  men diastolic blood pressure 5-number-summary
  women diastolic blood pressure 5-number-summary
  
  
```{r j,echo=FALSE}
men_sys_summary <- summary(SYSBP[GENDER==0])
men_sys_summary


women_sys_summary <- summary(SYSBP[GENDER==1])
women_sys_summary



men_dias_summary <- summary(DIASBP[GENDER==0])
men_dias_summary


women_dias_summary <- summary(DIASBP[GENDER==1])
women_dias_summary
```


k) Boxplot of men and women

Results are in that order
  men systolic blood pressure boxplot
# No outliers and the median of this data seen as about 120, minimum and maximum value seen about 90 and 148 respectively, and 1. quartile is about 110, 3. quartile is about 127.

  women systolic blood pressure boxplot
#There are outliers and median of this data seen as about 110, minimum and maximum value seen about 88 and 130 respectively, and 1. quartile is about 100, 3. quartile is about 120. Outliers are about 150 and 160.
  men diastolic blood pressure boxplot
#There are outliers and median of this data seen as about 77, minimum and maximum value seen about 57 and 95 respectively, and 1. quartile is about 70, 3. quartile is about 80. Outliers are about 97 and 100.
  women diastolic blood pressure boxplot
#There are outliers and median of this data seen as about 70, minimum and maximum value seen about 55 and 98 respectively, and 1. quartile is about 53, 3. quartile is about 78. Outlier is about 40.
  
  
```{r k}
boxplot(SYSBP[GENDER==0],main="MEN", xlab="Systolic Bp", horizontal = FALSE)

boxplot(SYSBP[GENDER==1],main="WOMEN", xlab="Systolic Bp", horizontal = FALSE)


boxplot(DIASBP[GENDER==0],main="MEN", xlab="Diastolic Bp", horizontal = FALSE)

boxplot(DIASBP[GENDER==1],main="WOMEN",  xlab="Daistolic Bp", horizontal = FALSE)
```


l) Stem-and-leaf plot of men and women

Results are in that order
  men systolic blood pressure Stem-and-leaf plot
  women systolic blood pressure Stem-and-leaf plot
  men diastolic blood pressure Stem-and-leaf plot
  women diastolic blood pressure Stem-and-leaf plot
  
  
```{r l}
stem(SYSBP[GENDER==0])

stem(SYSBP[GENDER==1])


stem(DIASBP[GENDER==0])

stem(DIASBP[GENDER==1])

```
m) Histograms of men and women

Results are in that order
  men systolic blood pressure histogram
  women systolic blood pressure histogram
  men diastolic blood pressure histogram
  women diastolic blood pressure histogram
  
  
```{r m}
hist(SYSBP[GENDER==0], right = FALSE, main = "MEN", xlab = "Systolic Bp")

hist(SYSBP[GENDER==1], right = FALSE, main = "WOMEN", xlab = "Systolic Bp")


hist(DIASBP[GENDER==0], right = FALSE, main = "MEN", xlab = "Diastolic Bp")

hist(DIASBP[GENDER==1], right = FALSE, main = "WOMEN", xlab = "Diastolic Bp")
```


m) Dotplots of men and women

Results are in that order
  men systolic blood pressure dotplot
  women systolic blood pressure dotplot
  men diastolic blood pressure dotplot
  women diastolic blood pressure dotplot
  
```{r n}

dotchart(SYSBP[GENDER==0], main = "MEN", xlab = "Sytolic Bp")

dotchart(SYSBP[GENDER==1], main = "WOMEN", xlab = "Sytolic Bp")

dotchart(DIASBP[GENDER==0], main = "MEN", xlab = "Diastolic Bp")

dotchart(DIASBP[GENDER==1], main = "WOMEN", xlab = "Diastolic Bp")

```



o) Tchebysheff’s Theorem

  Yes, we can use the Tchebysheff’s Theorem because of the usability any data set without a necessity like
mound shaped distibution or any kind. So the Tchebysheff’s Theorem also employable on this data set.



p) Empirical Rule

  The Empirical Rule just applicable on mound shaped data set. If we look this dataset, we realize that 
diastolic blood pressure of men and women are mound shaped, so emprical rule can be apply our data set.




q) Comperative Dotplots of men and women

Results are in that order
  men and women systolic blood pressure 
  men and women diastolic blood pressure 

```{r q}
dotchart(SYSBP[GENDER==0],SYSBP[GENDER==1], main = "MEN and WOMEN", xlab = "Systolic Bp")

dotchart(DIASBP[GENDER==0],DIASBP[GENDER==1], main = "MEN and WOMEN", xlab = "Diastolic Bp")
```


r) Relative frequency histograms of men and women

Results are in that order
  men systolic blood pressure relative freq. histogram
  women systolic blood pressure relative freq. histogram
  men diastolic blood pressure relative freq. histogram
  women diastolic blood pressure relative freq. histogram
  
```{r r}
# Skewed left
hist(SYSBP[GENDER==0], prob=TRUE, main = "MEN", xlab = "Systolic Bp")

# Skewed right
hist(SYSBP[GENDER==1], prob=TRUE, main = "WOMEN", xlab = "Systolic Bp")

# Mound shaped
hist(DIASBP[GENDER==0], prob=TRUE, main = "MEN", xlab = "Diastolic Bp")

# Mound shaped
hist(DIASBP[GENDER==1], prob=TRUE, main = "WOMEN", xlab = "Diastolic Bp")
```


s) Z-scores for largest and smallest observation of men and women

Results are in that order                                                                
  men systolic blood pressure maximum value z-score
# It may not be  outlier because it is close to 2
  
  men systolic blood pressure minimum value z-score
# Outlier because it is smaller than -3
  
  women systolic blood pressure maximum value z-score
# Outlier because it is greater than -3
  
  women systolic blood pressure minimum value z-score
# Not an outlier because it is between -2 and 2
  
  men diastolic blood pressure maximum value z-score
# It may not be  outlier because it is close to 2
  
  men diastolic blood pressure minimum value z-score
# Not an outlier because it is between -2 and 2

  women diastolic blood pressure maximum value z-score
# It can be say outlieri it is too close to -3

  women diastolic blood pressure minimum value z-score
# Outlier because it is smaller than -3


```{r s,echo=FALSE}
men_sys_zscoremax <- (men_sys_max - men_sys_mean)/men_sys_stddev
men_sys_zscoremax


men_sys_zscoremin <- (men_sys_min - men_sys_mean)/men_sys_stddev
men_sys_zscoremin


women_sys_zscoremax <- (women_sys_max - women_sys_mean)/women_sys_stddev
women_sys_zscoremax

women_sys_zscoremin <- (women_sys_min - women_sys_mean)/women_sys_stddev
women_sys_zscoremin

men_dias_zscoremax <- (men_dias_max - men_dias_mean)/men_dias_stddev
men_dias_zscoremax


men_dias_zscoremin <- (men_dias_min - men_dias_mean)/men_dias_stddev
men_dias_zscoremin

women_dias_zscoremax <- (women_dias_max - women_dias_mean)/women_dias_stddev
women_dias_zscoremax

women_dias_zscoremin <- (women_dias_min - women_dias_mean)/women_dias_stddev
women_dias_zscoremin
```


